---
title: "text frequency"
output: html_document
date: "2025-02-20"
---

```{r, warning = FALSE, message = FALSE}
library(readr)
library(dplyr)
library(lubridate)
library(ggplot2)
library(tidytext)
library(stringr)
library(textdata)
library(tidyr)
library(scales)
```


### Initital Data Cleaning: Pre-Process and Preparing Data
```{r}
# load the original data set 
full_data <- read.csv("/Users/avagrey/desktop/my_data.csv")

# clean the original data set
clean_data <- full_data |> 
  filter(!grepl("France|Sydney|London|Germany|Dubai|Israel|Ontario|Canada|Toronto|Thaiwan|China|UK|Norway|Europe|sydney|🇨🇦|England|India", author.location, ignore.case = TRUE)) |> #remove non-US countries
   # keep only tweets mentioning either candidate
  filter(grepl("Trump|Donald|Harris|Kamala", fullText, ignore.case = TRUE))|>
  filter(!grepl("-\\s*Kamala Harris", fullText)) |> # remove quotes by either canidate
  filter(!grepl("-\\s*Donald Trump", fullText)) |>
  mutate(fullText = tolower(fullText)) |>  #convert all text to lower case
  mutate(fullText = gsub("@\\w+", "", fullText)) |>  # remove mentions
  mutate(fullText = gsub("#\\w+", "", fullText)) |>  # remove hashtags
  mutate(fullText = gsub("[^[:alnum:]' ]", "", fullText)) |>
  mutate(fullText = gsub("\\s+", " ", fullText)) |>   
  distinct(fullText, author.id, .keep_all = TRUE) |> # remove duplicates 
  distinct(fullText, .keep_all = TRUE) 
```

### Further Cleaning to add month and candidate_mentioned columns
```{r}
# further cleaning to add months and specific candidates mentioned
valid_months <- month.abb  # built in abbreviation month string

clean_data <- clean_data |> 
  mutate(month = str_extract(createdAt, paste(valid_months, collapse = "|")))

# Define patterns to match variations of each candidate's name
trump_pattern <- regex("(trump|donald)", ignore_case = TRUE)
harris_pattern <- regex("(kamala|harris)", ignore_case = TRUE)

# Create a new column 'candidate_mentioned' based on the patterns
clean_data <- clean_data |>
  mutate(candidate_mentioned = case_when(
    str_detect(fullText, trump_pattern) & str_detect(fullText, harris_pattern) ~ "Trump, Harris",
    str_detect(fullText, trump_pattern) ~ "Trump",
    str_detect(fullText, harris_pattern) ~ "Harris"
  ))

#further cleaning to remove extra tweets not mentioning either candidate
clean_data <- clean_data |> filter(!is.na(candidate_mentioned))
```


### Understand aspects of the data
```{r}
z <- clean_data |> 
  group_by(candidate_mentioned) |> 
  summarize(n = n())

ggplot(z, aes(candidate_mentioned, n)) +  
  geom_col(show.legend = TRUE) +
  geom_text(aes(label = n), vjust = -0.5) +
  theme_minimal()
```

### Tokenize `fullText` to break up tweets into indivdal words to better understand frequency
```{r}
#tokenize the data
token_tweets <- clean_data |>
  unnest_tokens(word, fullText)
# count the individual word count with no restrictions
overall_count <- token_tweets |>
  count(word, sort = TRUE)

head(overall_count, 15)
```

### Visual of top words 
These make sense because a lot of search words were related to emotional starts like "i think..."
```{r}
token_tweets |> count(word, sort = TRUE)|>
  filter(n > 7900)|>
  mutate(word = reorder(word, n))|>
  ggplot(aes(word, n)) + 
  geom_col() +
  xlab(NULL) +
  coord_flip()
```

### First Step: remove basic lexicon stop words to see remaining word frequency
```{r}
# take 
stop_words2 <- stop_words |> filter(lexicon != "SMART")

token_tweets_rmstop <- token_tweets |>
  anti_join(stop_words)
```

### overall first looking at common words without stop words
```{r}
token_tweets_rmstop |> 
  count(word, sort = TRUE) |> 
  slice_max(n, n = 15) |>  
  mutate(word = reorder(word, n)) |> 
  ggplot(aes(word, n)) + 
  geom_col() + 
  xlab(NULL) + 
  coord_flip() +
  labs(title = "Top 15 Most Frequent Words in Tweets without Stop Words")

```

### look at it grouped by the different candidates
```{r}
# Count words by candidate mentioned
plot_stoprm_candidate <- token_tweets_rmstop |>  
  count(candidate_mentioned, word, sort = TRUE) |>  # Count words by candidate
  group_by(candidate_mentioned) |>  
  slice_max(n, n = 15) |>  # Select the top 10 words per candidate
  ungroup() |>  
  mutate(word = reorder_within(word, n, candidate_mentioned)) 

# Plot grouped word frequency by candidate
ggplot(plot_stoprm_candidate, aes(word, n, fill = candidate_mentioned)) +  
  geom_col(show.legend = TRUE) +  
  facet_wrap(~candidate_mentioned, scales = "free_y") +  
  xlab(NULL) +  
  coord_flip() +  
  scale_x_reordered() +  
  labs(title = "Most Common Words in Tweets by Candidate Mentioned",  
       y = "Word Count", fill = "Candidate Mentioned")
```

### look at it grouped by the different months
```{r}
# Count words by candidate mentioned
plot_stoprm_candidate <- token_tweets_rmstop |>  
  count(month, word, sort = TRUE) |>  # Count words by candidate
  group_by(month) |>  
  slice_max(n, n = 15) |>  # Select the top 10 words per candidate
  ungroup() |>  
  mutate(word = reorder_within(word, n, month),
         month = factor(month, c("Jul", "Aug", "Sep", "Oct", "Nov"))) 

# Plot grouped word frequency by candidate
ggplot(plot_stoprm_candidate, aes(word, n, fill = month)) +  # Add missing `+`
  geom_col(show.legend = TRUE) +  
  facet_wrap(~month, scales = "free_y") +  # Separate panels per candidate
  xlab(NULL) +  
  coord_flip() +  
  scale_x_reordered() +  # Ensure correct ordering of words
  labs(title = "Most Common Words in Tweets by Month",  
       y = "Word Count", fill = "Month")
```

### Now look at word counts with emotion related words
```{r}
# Load NRC sentiment lexicon
nrc_emotions <- get_sentiments("nrc") |>
  filter(word != "trump") |>
  filter(word != "ill") |>
  filter(word != "president")|>
  filter(word != "shell")

token_tweets_emotion <- token_tweets |> 
  inner_join(nrc_emotions, by = "word")

emotion_count <- token_tweets_emotion |>
  count(word, sort = TRUE)
```


### Overall frequency
```{r}
# now looking at emotions words specifically
token_tweets_emotion |>
  count(word, sort = TRUE) |> head(15)

token_tweets_emotion |> 
  count(word, sort = TRUE) |>  # Count occurrences of each word
  slice_max(n, n = 15) |>  # Select the top 15 most frequent words
  mutate(word = reorder(word, n)) |>  # Reorder for better visualization
  ggplot(aes(word, n)) + 
  geom_col() + 
  xlab(NULL) + 
  coord_flip() + 
  labs(title = "Top 15 Most Frequent Emotion Words in Tweets")
```

### Look at emotion frequency based by candidate mentions
```{r}
# Count emotion-related words by candidate mentioned
word_counts <- token_tweets_emotion |>  
  count(candidate_mentioned, word, sort = TRUE) |>  # Count words by candidate
  group_by(candidate_mentioned) |>  
  slice_max(n, n = 15) |>  # Select the top 10 words per candidate
  ungroup() |>  
  mutate(word = reorder_within(word, n, candidate_mentioned))  # Ensure proper ordering

# Plot grouped word frequency by candidate
ggplot(word_counts, aes(word, n, fill = candidate_mentioned)) +  
  geom_col(show.legend = TRUE) +  
  facet_wrap(~candidate_mentioned, scales = "free_y") +  # Separate panels per candidate
  xlab(NULL) +  
  coord_flip() +  
  scale_x_reordered() +  # Ensure correct ordering of words
  labs(title = "Most Common Emotion Words in Tweets by Candidate Mentioned",  
       y = "Word Count", fill = "Candidate Mentioned")

```

### Look at emotion frequency based by month 
```{r}
# Count emotion-related words by candidate mentioned
word_counts <- token_tweets_emotion |>  
  count(month, word, sort = TRUE) |>  # Count words by candidate
  group_by(month) |>  
  slice_max(n, n = 15) |>  # Select the top 10 words per candidate
  ungroup() |>  
  mutate(word = reorder_within(word, n, month))  # Ensure proper ordering

# Plot grouped word frequency by candidate
ggplot(word_counts, aes(word, n, fill = month)) +  
  geom_col(show.legend = TRUE) +  
  facet_wrap(~month, scales = "free_y") +  # Separate panels per candidate
  xlab(NULL) +  
  coord_flip() +  
  scale_x_reordered() +  # Ensure correct ordering of words
  labs(title = "Most Common Emotion Words in Tweets by Candidate Mentioned",  
       y = "Word Count", fill = "Candidate Mentioned")

```


# Changes overtime related to candidates (KAMALA)
```{r}
emotion_over_time <- function(df, candidate){
  word_count <- df |>
    filter(candidate_mentioned== candidate) |>
    count(month, word, sort = TRUE) |>
    group_by(month) |>
    slice_max(n, n = 15) |> # keep top 10 per month
    ungroup() |>
    mutate(word = reorder_within(word, n, month),
           month = factor(month, c("Jul", "Aug", "Sep", "Oct", "Nov")))
  
  ggplot(word_count, aes(word, n, fill = month)) +
    geom_col(show.legend = TRUE) +
    facet_wrap(~month, scales = "free_y") + # separate per candidate
    coord_flip() +
    scale_x_reordered() + # ensure correct order of words
    labs(title = paste("Most Common Emotion Words in Tweets for", candidate),
     y = "Word Count", fill = "Month") +
    theme(legend.position = "none")
}
```


```{r}
emotion_over_time(token_tweets_emotion, "Harris")
emotion_over_time(token_tweets_emotion, "Trump")
```


```{r}
emotion_candidate_overtime <- function(df, candidate, emotion) {
  candidate_clean <- df |>
    mutate(candidate_mentioned = tolower(candidate_mentioned),
           word = tolower(word))  # Ensure words are also in lowercase

  token_tweet_candidate <- candidate_clean |>
    filter(candidate_mentioned == tolower(candidate))

  token_tweet_candidate <- token_tweet_candidate |>
    mutate(month = factor(month, levels = month.abb))

  monthly_total <- token_tweet_candidate |>
    count(month)

  monthly_emotion <- token_tweet_candidate |>
    filter(word == tolower(emotion)) |>
    count(month, name = "emotion_count")

  monthly_data <- left_join(monthly_emotion, monthly_total, by = "month") |>
    mutate(references_per_tweet = emotion_count / n) |>
    arrange(month)  # Sort months properly

  ggplot(monthly_data, aes(x = month, y = references_per_tweet)) +
    geom_bar(stat = "identity", fill = "steelblue") +
    labs(title = paste("Normalized Frequency of '", emotion, "' Emotion Over Time for", candidate),
         x = "Month",
         y = paste("Proportion of Tweets Containing '", emotion, "'")) +
    theme_minimal()
}
```
```{r}
emotion_candidate_overtime(token_tweets_emotion, "Harris", "excited")
emotion_candidate_overtime(token_tweets_emotion, "Harris", "hate")
emotion_candidate_overtime(token_tweets_emotion, "Harris", "trust")
emotion_candidate_overtime(token_tweets_emotion, "Harris", "anger")

emotion_candidate_overtime(token_tweets_emotion, "Trump", "excited")
emotion_candidate_overtime(token_tweets_emotion, "Trump", "hate")
emotion_candidate_overtime(token_tweets_emotion, "Trump", "trust")
emotion_candidate_overtime(token_tweets_emotion, "Trump", "anger")
```






# looking at word overlap

```{r}
token_tweets_emotion_2 <- token_tweets_emotion |> filter(candidate_mentioned != "Trump, Harris")

frequency <- token_tweets_emotion_2 |> 
  count(candidate_mentioned, word) |> 
  group_by(candidate_mentioned) |> 
  mutate(proportion = n / sum(n)) |> 
  select(-n) |> 
  pivot_wider(names_from = candidate_mentioned, values_from = proportion, values_fill = 0)|>
  filter(Harris > 0 & Trump > 0)

# Plot with both Trump and Harris on a single scatter plot
ggplot(frequency, aes(x = Harris, y = Trump)) +
  geom_abline(color = "gray40", lty = 2) +  # Reference diagonal line
  geom_jitter(aes(color = abs(Harris - Trump)), alpha = 0.3, size = 2.5, width = 0.2, height = 0.2) + 
  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5, size = 3) +
  scale_x_log10(labels = percent_format()) +
  scale_y_log10(labels = percent_format()) +
  scale_color_gradient(low = "blue", high = "red") +  
  theme_minimal() +
  theme(legend.position = "none") +
  labs(
    title = "Word Proportions: Trump vs. Harris",
    x = "Harris",
    y = "Trump"
  )

```

```{r}
# Find words that appear only in Trump tweets (Harris proportion == 0)
exclusive_trump_words <- token_tweets_emotion_2 |> 
  count(candidate_mentioned, word) |> 
  group_by(candidate_mentioned) |> 
  mutate(proportion = n / sum(n)) |> 
  select(-n) |> 
  pivot_wider(names_from = candidate_mentioned, values_from = proportion, values_fill = 0)|>
  filter(Harris == 0 & Trump > 0) |>
  arrange(desc(Trump))

exclusive_harris_words <- token_tweets_emotion_2 |> 
  count(candidate_mentioned, word) |> 
  group_by(candidate_mentioned) |> 
  mutate(proportion = n / sum(n)) |> 
  select(-n) |> 
  pivot_wider(names_from = candidate_mentioned, values_from = proportion, values_fill = 0)|>
  filter(Harris > 0 & Trump == 0) |>
  arrange(desc(Harris))

# Display the exclusive words
exclusive_trump_words |> head(10)
exclusive_harris_words |> head(10)
```

```{r}
freq2 <- token_tweets_emotion_2 |> 
  count(candidate_mentioned, word) |> 
  group_by(candidate_mentioned) |> 
  mutate(proportion = n / sum(n)) |> 
  select(-n) |> 
  pivot_wider(names_from = candidate_mentioned, values_from = proportion, values_fill = 0)

cor.test(freq2$Harris, freq2$Trump)
```

