---
title: "lexicon exploration"
output: html_document
date: "2025-02-27"
---
```{r, warning = FALSE, message = FALSE}
library(readr)
library(dplyr)
library(lubridate)
library(ggplot2)
library(tidytext)
library(stringr)
library(textdata)
library(tidyr)
library(scales)
library(forcats)
library(wordcloud)
library(reshape2)
```


```{r}
# load the original data set 
full_data <- read.csv("my_data.csv")

# clean the original data set
clean_data <- full_data |> 
  filter(!grepl("France|Sydney|London|Germany|Dubai|Israel|Ontario|Canada|Toronto|Thaiwan|China|UK|Norway|Europe|sydney|🇨🇦|England|India", author.location, ignore.case = TRUE)) |> #remove non-US countries
  filter(grepl("Trump|Donald|Harris|Kamala", fullText, ignore.case = TRUE)) |>  # keep only tweets mentioning either candidate
  filter(!grepl("-\\s*Kamala Harris", fullText)) |> # remove quotes by either canidate
  filter(!grepl("-\\s*Donald Trump", fullText)) |>
  mutate(fullText = tolower(fullText)) |>  #convert all text to lower case
  mutate(fullText = gsub("@\\w+", "", fullText)) |>  # remove mentions
  mutate(fullText = gsub("#\\w+", "", fullText)) |>  # remove hashtags
  mutate(fullText = gsub("[^[:alnum:]' ]", "", fullText)) |>
  mutate(fullText = gsub("\\s+", " ", fullText)) |>   
  distinct(fullText, author.id, .keep_all = TRUE) |> # remove duplicates 
  distinct(fullText, .keep_all = TRUE) 
```


```{r}
# further cleaning to add months and specific candidates mentioned
valid_months <- month.abb  # built in abbreviation month string

clean_data <- clean_data |> 
  mutate(month = str_extract(createdAt, paste(valid_months, collapse = "|")))

# Define patterns to match variations of each candidate's name
trump_pattern <- regex("(trump|donald)", ignore_case = TRUE)
harris_pattern <- regex("(kamala|harris)", ignore_case = TRUE)

# Create a new column 'candidate_mentioned' based on the patterns
clean_data <- clean_data |>
  mutate(candidate_mentioned = case_when(
    str_detect(fullText, trump_pattern) & str_detect(fullText, harris_pattern) ~ "Trump, Harris",
    str_detect(fullText, trump_pattern) ~ "Trump",
    str_detect(fullText, harris_pattern) ~ "Harris"
  ))

#further cleaning to remove extra tweets not mentioning either candidate
clean_data <- clean_data |> filter(!is.na(candidate_mentioned))
```

```{r}
clean_data <- clean_data |> 
  mutate(day = day(as.POSIXct(createdAt, format="%a %b %d %H:%M:%S %z %Y", tz="UTC")))
```



```{r}
#tokenize the data
token_tweets <- clean_data |>
  unnest_tokens(word, fullText)
# count the individual word count with no restrictions
overall_count <- token_tweets |>
  count(word, sort = TRUE)

head(overall_count, 15) # again, just showing overall values
```

Thinking about potential lexicons - between either a politics specific lexicon or a more basic one like NRC.

Will need to think about:
- negations: could use a basic lexicon then add additional values + words

- good that the data is shorter to avoid potential zero-ing out 


```{r}
#Now, working with data to explore candidate specific trends
candidate_name <- c("Harris", "Trump", "Trump, Harris")

#Group my candidate to be able to analyze specific Trump or Harris Tweets and emotional trends
tidy_candidates <- clean_data |>
  group_by(candidate_mentioned) |>
  mutate(detection = cumsum(str_detect(fullText, str_c(candidate_name, collapse = "|")))) |>  # Fix parentheses
  ungroup() |>
  unnest_tokens(word, fullText)

months <- c("Jul", "Aug", "Sep", "Oct", "Nov")
tidy_month <- clean_data |>
  group_by(month) |>
  mutate(detection = cumsum(str_detect(fullText, str_c(months, collapse = "|")))) |>  # Fix parentheses
  ungroup() |>
  unnest_tokens(word, fullText)
```

```{r}
emotion_trends <- function(candidate, emotion){
  nrc_emoton <- get_sentiments("nrc")|>
    filter(sentiment == emotion)
  
  candidate_emotion <- tidy_candidates |>
    filter(candidate_mentioned == candidate)|>
    inner_join(nrc_emoton) |>
    count(word, sort = TRUE)
  
  return(candidate_emotion)
}

```


```{r}
#fear
emotion_trends("Trump", "fear")
emotion_trends("Harris", "fear")
```

```{r}
emotion_trends("Trump", "anger")
emotion_trends("Harris", "anger")
```

```{r}
emotion_trends("Trump", "disgust")
emotion_trends("Harris", "disgust")
```

```{r}
emotion_trends("Trump", "trust")
emotion_trends("Harris", "trust")
```

```{r}
emotion_trends("Trump", "joy")
emotion_trends("Harris", "joy")
```

```{r}
emotion_trends("Trump", "anticipation")
emotion_trends("Harris", "anticipation")
```

```{r}
#Now, use the tidy month data to analyze positive and negative sentiment trends for each candidate over the five months leading up to the election
trump_sentiment <- tidy_month |> 
  filter(candidate_mentioned == "Trump") |>  
  filter(!word %in% c("trump")) |>  # Exclude words before sentiment join
  inner_join(get_sentiments("bing"), by = "word") |>  
  count(month, day, sentiment) |>  
  spread(sentiment, n, fill = 0) |>  
  mutate(sentiment = positive - negative)

# Define correct month order
month_order <- c("Jul", "Aug", "Sep", "Oct", "Nov")

trump_sentiment <- trump_sentiment |> 
  mutate(month = factor(month, levels = month_order))  # Force chronological order



ggplot(trump_sentiment, aes(x = day, y = sentiment, fill = sentiment > 0)) + 
  geom_col(show.legend = FALSE, width = 0.7) +  # Fixed bar width for uniformity
  facet_wrap(~month, ncol = 2, scales = "free_x") +
  scale_fill_manual(values = c("red", "green"), labels = c("Negative", "Positive")) +  
  labs(x = "Day", y = "Sentiment Score", title = "TRUMP Sentiment Over Time") +
  theme_minimal()
```

```{r}
harris_sentiment <- tidy_month |> 
  filter(candidate_mentioned == "Harris") |>  # Keep only Trump-related data
  inner_join(get_sentiments("bing")) |>  
  count(month, day, sentiment) |>  
  spread(sentiment, n, fill = 0) |>  
  mutate(sentiment = positive - negative)


# Define correct month order
month_order <- c("Jul", "Aug", "Sep", "Oct", "Nov")

harris_sentiment <- harris_sentiment |> 
  mutate(month = factor(month, levels = month_order))  # Force chronological order



ggplot(harris_sentiment, aes(x = day, y = sentiment, fill = sentiment > 0)) + 
  geom_col(show.legend = FALSE, width = 0.7) +  # Fixed bar width for uniformity
  facet_wrap(~month, ncol = 2, scales = "free_x") +
  scale_fill_manual(values = c("red", "green"), labels = c("Negative", "Positive")) +  
  labs(x = "Day", y = "Sentiment Score", title = "HARRIS Sentiment Over Time") +
  theme_minimal()

```
# Need to be careful with interpreting these values s some words may not have negative sentiments about the candidate but are labeled as negative like assasniation 

I used bing here because it has a simpler "pos" "neg" model

Below I look at NRC words contribution and BING to see what words are driving these values for each candidate


```{r}
nrc_word_count_trump <- tidy_candidates |> 
  filter(word != "trump", word != "vote") |>  
  inner_join(get_sentiments("nrc"), by = "word") |>  
  count(candidate_mentioned, word, sentiment, sort = TRUE) |>  # Keep candidate_mentioned
  ungroup()|>
  filter(candidate_mentioned == "Trump")

nrc_word_count_harris <- tidy_candidates |> 
  filter(word != "trump", word != "vote") |>  
  inner_join(get_sentiments("nrc"), by = "word") |>  
  count(candidate_mentioned, word, sentiment, sort = TRUE) |>  # Keep candidate_mentioned
  ungroup()|>
  filter(candidate_mentioned == "Harris")

nrc_word_count_trump
nrc_word_count_harris
```



```{r}
nrc_word_count_trump |>
  group_by(sentiment) |>
  top_n(10) |>
ungroup()|>
  mutate(word = reorder(word, n)) |>
  ggplot(aes(word, n, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, scales = "free_y") +
  labs(y = "Contribution to sentiment TRUMP NRC",
       x = NULL) +
  coord_flip()

nrc_word_count_harris |>
  group_by(sentiment) |>
  top_n(10) |>
ungroup()|>
  mutate(word = reorder(word, n)) |>
  ggplot(aes(word, n, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, scales = "free_y") +
  labs(y = "Contribution to sentiment HARRIS NRC",
       x = NULL) +
  coord_flip()

bing_word_count_trump |>
  group_by(sentiment) |>
  top_n(10) |>
ungroup()|>
  mutate(word = reorder(word, n)) |>
  ggplot(aes(word, n, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, scales = "free_y") +
  labs(y = "Contribution to sentiment TRUMP BING",
       x = NULL) +
  coord_flip()

bing_word_count_harris |>
  group_by(sentiment) |>
  top_n(10) |>
ungroup()|>
  mutate(word = reorder(word, n)) |>
  ggplot(aes(word, n, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, scales = "free_y") +
  labs(y = "Contribution to sentiment HARRIS BING",
       x = NULL) +
  coord_flip()
```

```{r}
word_cloud_candidate <- function(candidate, sentiment){
  tidy_candidate <- tidy_candidates |>
    filter(candidate_mentioned == candidate)|>
    filter(word != "trump")|>
    inner_join(get_sentiments(sentiment)) |>
    count(word, sentiment, sort = TRUE) |>
    acast(word ~ sentiment, value.var = "n", fill = 0) |>
    comparison.cloud(colors = c("darkred", "forestgreen"), max.words = 50)
}

word_cloud_candidate("Trump", "bing")
word_cloud_candidate("Harris", "bing")
```

```{r}
election_sentences <- clean_data |>
  unnest_tokens(sentence, fullText, token = "sentences")

bingsentiment <- get_sentiments("bing")|>
    filter(sentiment == "positive")
  
word_count <- tidy_candidates |>
    group_by(candidate_mentioned, month)|>
    summarize(words = n())
  
tidy_candidates |> 
    semi_join(bingsentiment)|>
    group_by(candidate_mentioned, month)|>
    summarize(sentimentwords = n())|>
    left_join(word_count, by = c("candidate_mentioned", "month")) |>
    mutate(ratio = sentimentwords/words) |>
    filter(candidate_mentioned != "Trump, Harris")|>
    top_n(1)|>
    ungroup()


```

